---
title: "generating"
author: "Laura Jochim"
date: "2025-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages, message=FALSE, warning=FALSE}
#load packages

library(tidyverse)
library(rvinecopulib)
library(kde1d)
library(reticulate)
library(synthpop)
library(utils)
library(pryr)
library(dplyr)
library(progressr)
library(gplots)
library(glmnet)
library(purrr)
library(caret)
library(ggplot2)
library(philentropy)
library(progress)
library(kableExtra)

#set virtual environment
reticulate::use_condaenv("thesis", required = T) 
```

This file aims to have a first exploration stage of the EHR from Figshare. We will dive into synthesizing it using three methods, namely: synthpop, CtGAN, and vine copulas. Lastly, we will dive into the evaluation of the synthetic data, and how well these evaluation methods tell us how good the synthetic data is. In this project we will only focus on the utility of the synthetic data, privacy which is also something to evaluate the synthetic data on will not be part of the evaluation. 


```{r}
# load data
ehr <- read.csv("../data/clean/ehr.csv")[, -1]
```


Now we have a nicely cleaned data, thank you Said, to work with. Each row represents an individual and each column a variable the patient was measured on, e.g. age, race, red blood cell count, white blood cell count, platelet count etc... 
It is important to note yet again that also this electronic health care record of 10.000 parients is a syntesized and not real one. The following step will inlcude synthesizing data based on this EHR.

## Synthesizing data
In this step I will dive into the process of synthesizing data using three methods, `synthpop`, `vinecopula`, and `ctGAN`. 

### Synthpop
First I will start with the synthpop method: (describe a bit what synthpop does)


```{r synthpop, message= F, warning=FALSE}

#set seed for reproducabulity
set.seed(123)


#change characters into factors
ehr_fac <- ehr %>%
  mutate(across(where(is.character), as.factor))

#create synthetic data using synthpop
  #record time
start_time <- Sys.time()

ehr_synthpop <- synthpop::syn(ehr_fac)

  #record end time
end_time <- Sys.time()
# Calculate the difference in time
execution_time <- end_time - start_time
print(execution_time)
#it took around 40 seconds to run synthpop

#get a bit of an idea of the data
summary(ehr_synthpop)
View(ehr_synthpop$syn)
#GGally::ggpairs(data.frame(ehr_synthpop$syn))

```
Successful synthetic data obtained using synthpop:) Unfortunately for now the birthdate is removed. Above there are a plots comparing the frequencies of the values of each variable for the original and generated data along with the standardized propensity score measure. The plots show that the original and synthetic data look rather real. 


### Vine Copula
```{r vine copula 1, message=F, warning=FALSE}
source("https://raw.githubusercontent.com/selbouhaddani/copula_vps/refs/heads/main/scripts/functions/estimate_vinecopula_from_data.R")

#check the levels of the categorical variables

levels(ehr_fac$PatientGender) 
levels(ehr_fac$PatientLanguage)  
levels(ehr_fac$PatientMaritalStatus) 
levels(ehr_fac$PatientRace)  
levels(ehr_fac$outcome)

#
ehr_fac <- ehr %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(CBC_BASOPHILS = as.factor(CBC_BASOPHILS))
ehr_fac <- ehr_fac %>% 
  mutate(across(where(is.factor), \(x) factor(x, levels = x %>% table %>% sort(decreasing = TRUE) %>% names)))


# re-code factors into numerical variables
ehr_num <- ehr_fac %>%
  mutate(
    outcome = as.numeric(outcome),  
    PatientGender = as.numeric(PatientGender),  
    PatientLanguage = as.numeric(PatientLanguage),  
    PatientMaritalStatus = as.numeric(PatientMaritalStatus),  
    PatientRace = as.numeric(PatientRace),
    CBC_BASOPHILS = as.numeric(CBC_BASOPHILS)
  )



#transform tibble to data frame
#ehr_4vc <- as.data.frame(ehr_4vc)

# define discrete and continuous variables
var_types <- c(
  "d",  # outcome (factor)
  "c",  # PatientAge (numeric)
  "d",  # PatientGender (factor)
  "d",  # PatientLanguage (factor)
  "c",  # PatientLOS (numeric)
  "d",  # PatientMaritalStatus (factor)
  "c",  # PatientPopulationPercentageBelowPoverty (numeric)
  "d",  # PatientRace (factor)
  "c",  # CBC_ABSOLUTE_LYMPHOCYTES (numeric)
  "c",  # CBC_ABSOLUTE_NEUTROPHILS (numeric)
  "d",  # CBC_BASOPHILS (factor)
  "c",  # CBC_EOSINOPHILS (numeric)
  "c",  # CBC_HEMATOCRIT (numeric)
  "c",  # CBC_HEMOGLOBIN (numeric)
  "c",  # CBC_LYMPHOCYTES (numeric)
  "c",  # CBC_MCH (numeric)
  "c",  # CBC_MCHC (numeric)
  "c",  # CBC_MEAN_CORPUSCULAR_VOLUME (numeric)
  "c",  # CBC_MONOCYTES (numeric)
  "c",  # CBC_NEUTROPHILS (numeric)
  "c",  # CBC_PLATELET_COUNT (numeric)
  "c",  # CBC_RDW (numeric)
  "c",  # CBC_RED_BLOOD_CELL_COUNT (numeric)
  "c",  # CBC_WHITE_BLOOD_CELL_COUNT (numeric)
  "c",  # METABOLIC_ALBUMIN (numeric)
  "c",  # METABOLIC_ALK_PHOS (numeric)
  "c",  # METABOLIC_ALT_SGPT (numeric)
  "c",  # METABOLIC_ANION_GAP (numeric)
  "c",  # METABOLIC_AST_SGOT (numeric)
  "c",  # METABOLIC_BILI_TOTAL (numeric)
  "c",  # METABOLIC_BUN (numeric)
  "c",  # METABOLIC_CALCIUM (numeric)
  "c",  # METABOLIC_CARBON_DIOXIDE (numeric)
  "c",  # METABOLIC_CHLORIDE (numeric)
  "c",  # METABOLIC_CREATININE (numeric)
  "c",  # METABOLIC_GLUCOSE (numeric)
  "c",  # METABOLIC_POTASSIUM (numeric)
  "c",  # METABOLIC_SODIUM (numeric)
  "c",  # METABOLIC_TOTAL_PROTEIN (numeric)
  "c",  # URINALYSIS_PH (numeric)
  "c",  # URINALYSIS_RED_BLOOD_CELLS (numeric)
  "c"   # URINALYSIS_WHITE_BLOOD_CELLS (numeric)
)
```


```{r vine copula 2, message=F, warning=FALSE}
set.seed(123)

#record time
start_time <- Sys.time()  

cat("Starting...\n")
print(start_time)
 system.time({fit_vc <- estimate_vinecopula_from_data(ehr_num[1:500,], var_types = var_types)})
## 25 min
cat("Ending...\n")
print(Sys.time() - start_time)
print(Sys.time())

N <- nrow(ehr_num)

#load("fit_vc_9500_20250325.RData")

ehr_vc <- simulate.estVineCopula(fit_vc, 1.5*N)
```

```{r look at vine copula, message=F, warning=FALSE}
#contour plots of the pair copulas
contour(fit_vc$vine_copula)

#the R-vine diagram
plot(fit_vc$vine_copula)

#count of which density was used how many times to estimate the joint distribution
table(unlist(lapply(fit_vc$vine_copula$pair_copulas, function(inner_list) {
  sapply(inner_list, function(x) x$family)
})))


```


```{r vine copula 3, message=F, warning=FALSE}


#transform the categorical columns back the original categories
ehr_vc <- ehr_vc %>% filter(PatientGender %in% unique(ehr_num$PatientGender)) %>%
  mutate(PatientGender = factor(PatientGender, labels = levels(ehr_fac$PatientGender)))

ehr_vc <- ehr_vc %>% filter(PatientLanguage %in% unique(ehr_num$PatientLanguage)) %>%
  mutate(PatientLanguage = factor(PatientLanguage, labels = levels(ehr_fac$PatientLanguage)))

ehr_vc <- ehr_vc %>% filter(PatientMaritalStatus %in% unique(ehr_num$PatientMaritalStatus)) %>%
  mutate(PatientMaritalStatus = factor(PatientMaritalStatus, labels = levels(ehr_fac$PatientMaritalStatus)))

ehr_vc <- ehr_vc %>% filter(PatientRace %in% unique(ehr_num$PatientRace)) %>%
  mutate(PatientRace = factor(PatientRace, labels = levels(ehr_fac$PatientRace)))

ehr_vc <- ehr_vc %>% filter(outcome %in% unique(ehr_num$outcome)) %>%
  mutate(outcome = factor(outcome, labels = levels(ehr_fac$outcome)))

ehr_vc <- ehr_vc %>% filter(CBC_BASOPHILS %in% unique(ehr_num$CBC_BASOPHILS)) %>%
  mutate(CBC_BASOPHILS = factor(CBC_BASOPHILS, labels = levels(ehr_fac$CBC_BASOPHILS)))

ehr_vc <- ehr_vc[1:N,]
dim(ehr_vc)


```

## vine copula tables
```{r vine copula tables for EHR}
# 1. get a summary data.frame of all pair‐copulas
sum_df <- summary(fit_vc$vine_copula)

library(dplyr)
library(tidyr)
library(purrr)
library(stringr)

# 1) First, build a character vector of your variable names in index order:
var_names <- c(
  "outcome",
  "PatientAge",
  "PatientGender",
  "PatientLanguage",
  "PatientLOS",
  "PatientMaritalStatus",
  "PatientPopulationPercentageBelowPoverty",
  "PatientRace",
  "CBC_ABSOLUTE_LYMPHOCYTES",
  "CBC_ABSOLUTE_NEUTROPHILS",
  "CBC_BASOPHILS",
  "CBC_EOSINOPHILS",
  "CBC_HEMATOCRIT",
  "CBC_HEMOGLOBIN",
  "CBC_LYMPHOCYTES",
  "CBC_MCH",
  "CBC_MCHC",
  "CBC_MEAN_CORPUSCULAR_VOLUME",
  "CBC_MONOCYTES",
  "CBC_NEUTROPHILS",
  "CBC_PLATELET_COUNT",
  "CBC_RDW",
  "CBC_RED_BLOOD_CELL_COUNT",
  "CBC_WHITE_BLOOD_CELL_COUNT",
  "METABOLIC_ALBUMIN",
  "METABOLIC_ALK_PHOS",
  "METABOLIC_ALT_SGPT",
  "METABOLIC_ANION_GAP",
  "METABOLIC_AST_SGOT",
  "METABOLIC_BILI_TOTAL",
  "METABOLIC_BUN",
  "METABOLIC_CALCIUM",
  "METABOLIC_CARBON_DIOXIDE",
  "METABOLIC_CHLORIDE",
  "METABOLIC_CREATININE",
  "METABOLIC_GLUCOSE",
  "METABOLIC_POTASSIUM",
  "METABOLIC_SODIUM",
  "METABOLIC_TOTAL_PROTEIN",
  "URINALYSIS_PH",
  "URINALYSIS_RED_BLOOD_CELLS",
  "URINALYSIS_WHITE_BLOOD_CELLS",
  "outcome_d",
  "PatientGender_d",
  "PatientLanguage_d",
  "PatientMaritalStatus_d",
  "PatientRace_d",
  "CBC_BASOPHILS_d"
)

# 2) grab summary and coerce to plain df
sum_df <- as.data.frame(summary(fit_vc$vine_copula))

flat_tbl <- sum_df %>%
  # strip off 'c(', ')', and all double-quotes
  mutate(
    conditioned_clean = str_remove_all(conditioned, 'c\\(|\\)|"')
  ) %>%
  # now separate on the comma into two integer columns
  separate(
    conditioned_clean,
    into    = c("v1_idx", "v2_idx"),
    sep     = ",\\s*",
    convert = TRUE
  ) %>%
  # map indices → names
  mutate(
    var1    = var_names[v1_idx],
    var2    = var_names[v2_idx],
    # build the conditioned-on string as before
    cond_on = map_chr(
      conditioning,
      ~ if (length(.x)==0) "" else paste(var_names[.x], collapse = ", ")
    )
   ) %>%
  select(
    tree, edge,
    var1, var2, cond_on,
    family, rotation, parameters
  )

print(flat_tbl)

tree1_tbl <- flat_tbl %>% filter(tree == 1) %>%
  select(var1, var2, family, rotation, parameters)

```

### ctGAN
```{r ctGAN 1, warning=FALSE, message=FALSE}
ctGAN <- import("ctgan")


discrete_columns <- c('outcome', 'PatientGender', 'PatientLanguage', 'PatientMaritalStatus', 'PatientRace', 'CBC_BASOPHILS')

# Ensure categorical variables are characters (not factors)
#ehr[discrete_columns] <- lapply(ehr[discrete_columns], as.character)

# Convert the data to a Pandas DataFrame
#ehr_py <- r_to_py(ehr)


```


```{r ctGAN 2, warning=FALSE, message=FALSE}

set.seed(123)

#record time
start_time <- Sys.time()

ctgan_700 <- ctGAN$CTGAN(epochs = as.integer(700)) # too few epochs, but with 50 it takes around 3 minutes ~ so I expect 500 to take around 30 min

#default hyperparameters:
#batch_size: 500
#generator_dim: [256, 256, 256]
#discriminator_dim: [256, 256, 256]
#embedding_dim: 128
#generator_lr: 2e-4
#discriminator_lr: 2e-4


ctgan_700$fit(ehr, discrete_columns)
ehr_ctgan_700 <- ctgan_700$sample(9662L)

  #record end time
end_time <- Sys.time()
# Calculate the difference in time
execution_time <- end_time - start_time
print(execution_time)

#It took around 2.5 hours to run 
ctgan_700$save("ctgan_model.pkl")
#ctgan_bc <- ctgan$CTGANSynthesizer.load("ctgan_model.pkl")

ctgan_700$loss_values
################
#lil comment to myself - CTGAN runs when only doing the basic pre processing of the data - make sure vinecopula's dat is different to not mess with it
# dont transform characters to factors or bring them back to characters 
```

```{r plot}


########################################

# Convert loss values to a data frame
loss_df <- as.data.frame(ctgan_700$loss_values)

# Convert columns to numeric (if necessary)
loss_df$Epoch <- as.numeric(loss_df$Epoch)
loss_df$`Generator Loss` <- as.numeric(loss_df$`Generator Loss`)
loss_df$`Discriminator Loss` <- as.numeric(loss_df$`Discriminator Loss`)

# Plot Generator and Discriminator Loss
ggplot(loss_df, aes(x = Epoch)) +
  geom_line(aes(y = `Generator Loss`, color = "Generator Loss")) +
  geom_line(aes(y = `Discriminator Loss`, color = "Discriminator Loss")) +
  labs(title = "CTGAN Training Loss Over Epochs",
       x = "Epoch",
       y = "Loss",
       color = "Loss Type") +
  theme_minimal()

```
#####################################################################

```{r hyperparameter-tuning-ctgan_ehr, message=FALSE, warning=FALSE}
# Set the seed for reproducibility
set.seed(123)

#------------------------------
# Load Required Libraries
#------------------------------
library(future)
library(furrr)
library(dplyr)
library(philentropy)
library(ggplot2)

# Import Python CTGAN library
ctgan <- import("ctgan")
np <- import("numpy")
py_builtin <- import("builtins")

# Set up parallel backend using multisession with available cores.
plan(multisession, workers = availableCores())


# Specify the six categorical columns.
discrete_columns_ehr <- c('outcome', 
                          'PatientGender', 
                          'PatientLanguage', 
                          'PatientMaritalStatus', 
                          'PatientRace', 
                          'CBC_BASOPHILS')


# Automatically identify numeric columns (all columns not in discrete_columns_ehr)
numeric_columns_ehr <- setdiff(names(ehr), discrete_columns_ehr)

# Convert data to Python object (will re-create within each worker later; keeping it here for reference)
py_ehr <- r_to_py(ehr)

#------------------------------
# Helper Functions (Renamed with _ehr)
#------------------------------
compute_tvd <- function(p, q) {
  mat <- rbind(p, q)
  return(as.numeric(distance(mat, method = "manhattan") / 2))
}

evaluate_tvd_cat <- function(real_vec, synth_vec) {
  real_tab <- prop.table(table(real_vec))
  synth_tab <- prop.table(table(synth_vec))
  all_levels <- union(names(real_tab), names(synth_tab))
  real_probs <- sapply(all_levels, function(x) ifelse(x %in% names(real_tab), real_tab[x], 0))
  synth_probs <- sapply(all_levels, function(x) ifelse(x %in% names(synth_tab), synth_tab[x], 0))
  tvd <- compute_tvd(real_probs, synth_probs)
  return(tvd)
}

evaluate_ks <- function(real_vec, synth_vec) {
  ks_stat <- ks.test(real_vec, synth_vec)$statistic
  return(as.numeric(ks_stat))
}

evaluate_quality <- function(real_df, synth_df, numeric_cols, discrete_cols) {
  ks_vals <- sapply(numeric_cols, function(col) {
    evaluate_ks(real_df[[col]], synth_df[[col]])
  })
  mean_ks <- mean(ks_vals)
  
  tvd_vals <- sapply(discrete_cols, function(col) {
    evaluate_tvd_cat(real_df[[col]], synth_df[[col]])
  })
  mean_tvd <- mean(tvd_vals)
  
  overall_metric <- (mean_ks + mean_tvd) / 2
  return(overall_metric)
}


#------------------------------
# Hyperparameter Tuning Setup (Create Parameter Grid)
#------------------------------
epochs_values <- c(700)
batch_size_values <- c(100, 200)
generator_lr_values <- c(0.0001, 0.001)
discriminator_lr_values <- c(0.0001,0.001)

generator_dim_options <- list(c(64, 64), c(128, 64))
discriminator_dim_options <- list(c(64, 64), c(128, 64))
embedding_dim_options <- c(64)

# Compute total iterations 
total_iter <- length(epochs_values) * length(batch_size_values) *
  length(generator_lr_values) * length(discriminator_lr_values) *
  length(generator_dim_options) * length(discriminator_dim_options) *
  length(embedding_dim_options)

# Instead of nested loops, create a parameter grid data frame.
# (Here we use expand.grid for scalars and then add the list-valued dimensions.)

param_grid <- expand.grid(
  epochs = epochs_values,
  batch_size = batch_size_values,
  generator_lr = generator_lr_values,
  discriminator_lr = discriminator_lr_values,
  embedding_dim = embedding_dim_options,
  stringsAsFactors = FALSE
)

# Expand grid for generator_dim (list column)
param_grid$generator_dim <- list(generator_dim_options[[1]])
param_grid <- param_grid[rep(seq_len(nrow(param_grid)), each = length(generator_dim_options)), ]
param_grid$generator_dim <- rep(generator_dim_options, times = nrow(param_grid) / length(generator_dim_options))

# Expand grid for discriminator_dim (list column)
param_grid$discriminator_dim <- discriminator_dim_options[[1]]
param_grid <- param_grid[rep(seq_len(nrow(param_grid)), each = length(discriminator_dim_options)), ]
param_grid$discriminator_dim <- rep(discriminator_dim_options, times = nrow(param_grid) / length(discriminator_dim_options))

# ------------------------------

#------------------------------
# Parallel Processing Using future_map
#------------------------------

process_config_ehr <- function(i) {
  # Import Python modules in this worker.
  ctgan <- import("ctgan")
  py_builtin <- import_builtins()
  np <- import("numpy")
  
  # Extract parameters from the i-th row of param_grid_ehr.
  row <- param_grid[i, ]
  epochs <- row$epochs
  batch_size <- row$batch_size
  g_lr <- row$generator_lr
  d_lr <- row$discriminator_lr
  emb_dim <- row$embedding_dim
  
  gen_dim <- row$generator_dim[[1]]
  disc_dim <- row$discriminator_dim[[1]]
  
  # Re-create the Python object for the data in this worker.
  py_ehr <- r_to_py(ehr)
  
  # Convert dimensions to Python tuples.
  py_gen_dim <- py_builtin$tuple(as.integer(gen_dim))
  py_disc_dim <- py_builtin$tuple(as.integer(disc_dim))
  
  # Create the CTGAN synthesizer.
  synthesizer <- ctgan$CTGAN(
    epochs = as.integer(epochs),
    batch_size = as.integer(batch_size),
    generator_lr = g_lr,
    discriminator_lr = d_lr,
    generator_dim = py_gen_dim,
    discriminator_dim = py_disc_dim,
    embedding_dim = as.integer(emb_dim)
  )
  
  synthesizer$fit(ehr, discrete_columns_ehr)
  
  # Save the loss plot.
  loss_df_ehr <- as.data.frame(synthesizer$loss_values)
  loss_df_ehr$Epoch <- as.numeric(loss_df_ehr$Epoch)
  loss_df_ehr$`Generator Loss` <- as.numeric(loss_df_ehr$`Generator Loss`)
  loss_df_ehr$`Discriminator Loss` <- as.numeric(loss_df_ehr$`Discriminator Loss`)
  plot_filename <- sprintf("ehr_ctgan_loss_epochs%d_batch%d_genlr%s_disclr%s_gendim%s_discdim%s_embdim%d.png",
                           epochs, batch_size, g_lr, d_lr,
                           paste(gen_dim, collapse = "-"),
                           paste(disc_dim, collapse = "-"),
                           emb_dim)
  p_ehr <- ggplot(loss_df_ehr, aes(x = Epoch)) +
    geom_line(aes(y = `Generator Loss`, color = "Generator Loss")) +
    geom_line(aes(y = `Discriminator Loss`, color = "Discriminator Loss")) +
    labs(title = "CTGAN Training Loss Over Epochs",
         x = "Epoch", y = "Loss", color = "Loss Type") +
    theme_minimal() +
    theme(plot.background = element_rect(fill = "white", color = NA))
  ggsave(filename = plot_filename, plot = p_ehr, width = 8, height = 5, bg = "white")
  
  # Sample synthetic data.
  synth_py_ehr <- synthesizer$sample(as.integer(nrow(ehr)))
  
  # Evaluate quality.
  metric_val <- evaluate_quality(ehr, synth_py_ehr, numeric_columns_ehr, discrete_columns_ehr)
  
  # Return a data frame row.
  data.frame(
    epochs = epochs,
    batch_size = batch_size,
    generator_lr = g_lr,
    discriminator_lr = d_lr,
    generator_dim = paste(gen_dim, collapse = "-"),
    discriminator_dim = paste(disc_dim, collapse = "-"),
    embedding_dim = emb_dim,
    quality_metric = metric_val,
    loss_plot_file = plot_filename,
    stringsAsFactors = FALSE
  )
}

# Record start time.
start_time <- Sys.time()

# Process each configuration in parallel.
results_list_ehr <- future_map(1:nrow(param_grid), process_config_ehr, .options = furrr_options(seed = TRUE))
results_ehr <- bind_rows(results_list_ehr)

# Record end time and compute execution time.
end_time <- Sys.time()
execution_time <- end_time - start_time
print(execution_time)
print(results_ehr)

# Identify the best configuration (lowest quality_metric).
best_params_ehr <- results_ehr %>% filter(quality_metric == min(quality_metric))
cat("Best hyperparameters (allowed set):\n")
print(best_params_ehr)
```
```{r}
# look at only "converged" models:
# Define the converged model file names (make sure the names include the ".png" extension
# as they are saved by your code).
converged_files <- c(
  "ehr_ctgan_loss_epochs700_batch200_genlr1e-04_disclr1e-04_gendim128-64_discdim128-64_embdim64.png",
  "ehr_ctgan_loss_epochs700_batch200_genlr1e-04_disclr1e-04_gendim128-64_discdim64-64_embdim64.png",
  "ehr_ctgan_loss_epochs700_batch200_genlr1e-04_disclr1e-04_gendim64-64_discdim128-64_embdim64.png",
  "ehr_ctgan_loss_epochs700_batch200_genlr0.001_disclr1e-04_gendim64-64_discdim128-64_embdim64.png",
  "ehr_ctgan_loss_epochs700_batch200_genlr0.001_disclr1e-04_gendim64-64_discdim64-64_embdim64.png"
)
#this is actually too broad, only two models ended up converging: "ehr_ctgan_loss_epochs700_batch200_genlr0.001_disclr1e-04_gendim64-64_discdim128-64_embdim64.png"; "ehr_ctgan_loss_epochs700_batch200_genlr0.001_disclr1e-04_gendim64-64_discdim64-64_embdim64.png"

# Filter the results data frame to show only the rows where the loss_plot_file column is in converged_files.
converged_results <- results_ehr %>% filter(loss_plot_file %in% converged_files)

# Print the converged results
print(converged_results)

```


```{r best parameter}

set.seed(123)

#------------------------------
# Re-run Best Model to Save Synthetic Data
#------------------------------

# Specify the six categorical columns.
discrete_columns_ehr <- c('outcome', 
                          'PatientGender', 
                          'PatientLanguage', 
                          'PatientMaritalStatus', 
                          'PatientRace', 
                          'CBC_BASOPHILS')


best_epochs <- 700
best_batch_size <- 200
best_generator_lr <- 0.001
best_discriminator_lr <- 1e-04
best_embedding_dim <- 64

# Define the best network dimensions.
best_generator_dim <- c(64, 64)
best_discriminator_dim <- c(128, 64)

# Convert the network dimension vectors to Python tuples.
py_best_generator_dim <- py_builtin$tuple(as.integer(best_generator_dim))
py_best_discriminator_dim <- py_builtin$tuple(as.integer(best_discriminator_dim))

# Re-create the Python object for your data.
py_ehr <- r_to_py(ehr)

# Initialize and train the final CTGAN model with the best hyperparameters.
best_synthesizer_ehr <- ctgan$CTGAN(
  epochs = as.integer(best_epochs),
  batch_size = as.integer(best_batch_size),
  generator_lr = best_generator_lr,
  discriminator_lr = best_discriminator_lr,
  generator_dim = py_best_generator_dim,
  discriminator_dim = py_best_discriminator_dim,
  embedding_dim = as.integer(best_embedding_dim)
)

# Train the model (using your defined discrete columns).
best_synthesizer_ehr$fit(py_ehr, discrete_columns_ehr)

# Sample synthetic data: generate as many rows as in your original data.
ehr_ctgan_best <- best_synthesizer_ehr$sample(as.integer(nrow(ehr)))

#loss plot

# Convert loss values to a data frame
loss_df <- as.data.frame(best_synthesizer_ehr$loss_values)

# Convert columns to numeric (if necessary)
loss_df$Epoch <- as.numeric(loss_df$Epoch)
loss_df$`Generator Loss` <- as.numeric(loss_df$`Generator Loss`)
loss_df$`Discriminator Loss` <- as.numeric(loss_df$`Discriminator Loss`)

# Plot Generator and Discriminator Loss
ggplot(loss_df, aes(x = Epoch)) +
  geom_line(aes(y = `Generator Loss`, color = "Generator Loss")) +
  geom_line(aes(y = `Discriminator Loss`, color = "Discriminator Loss")) +
  labs(title = "CTGAN Training Loss Over Epochs",
       x = "Epoch",
       y = "Loss",
       color = "Loss Type") 
  theme_minimal()
``` 
